Optimal updating magnitude in adaptive flat-distribution-sampling simulations

### 2017-09-19 doc/invt.tex
change sig 0.29 to 0.28

\begin{table}[h]
  \caption{\label{tab:lj_error}
    Errors of the bias potential under
    the single-bin,
    Gaussian (with width $\sigma = 0.29$),
    and homogeneous bandpass (with mode cutoff $K = 26$)
    updating schemes
    on the LJ system.
    %
    The numbers have been averaged over multiple runs,
    and the predicted values are shown in parentheses for comparison;
    $\Err_\mathrm{init.}$ and $\Err_\mathrm{final}$
    denote the errors at the beginning and end of the production phase, respectively.
    %
    \note{The reference values can be found
    on the second line of \texttt{verr.log}.
    To check use \texttt{getav.py} under \texttt{prog/lj},
    which will show the simulation and prediction values.}%
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Schedule &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    %\hline
    %\multicolumn{4}{c}{
    %  $\varrho = 0.1$,
    %  $500$ bins,
    %  $\Gamma_1 \approx 111$,
    %  $\Gamma \approx 1.1\times10^3$.
    %} \\
    \hline
    Single-bin & $1/t$
    & $5.7(4.4)\times10^{-2}$
    & $9.1(8.8)\times10^{-6}$
    \\
    %\hline
    Gaussian & $1/t$
    & $6.5(5.9)\times10^{-3}$
    & $1.4(1.4)\times10^{-5}$
    \\
    %\hline
    Gaussian & optimal
    & $6.0(5.9)\times10^{-3}$
    & $2.6(2.6)\times10^{-6}$
    \\
    %\hline
    Bandpass & $1/t$
    & $8.0(7.6)\times10^{-3}$
    & $1.7(1.6)\times10^{-6}$
    \\
    \hline
  \end{tabular}
\end{table}

As shown in Fig. \ref{fig:lj_alpha},
the optimal schedule for the Gaussian updating scheme
with width $\sigma = 0.29$
was more complex than the inverse-time one.
%
We chose the Gaussian updating scheme here because
the optimal and inverse-time schedules were identical
in the single-bin and bandpass updating schemes.
%
The optimal schedule lay above the inverse-time schedule,
and its initial decay was roughly exponential,
in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
%
Since our equilibration phase was sufficiently long,
the initial updating magnitude, $\alpha(0)$,
was around the ideal value of $a_0/2 = 5\times10^{-5}$
[cf. Eq. \eqref{eq:half_alpha0}].

@@

For the Gaussian updating scheme,
the error depended not only on the schedule
but also on the width of the Gaussian.
%
Although widening the Gaussian
can suppress local noise,
it can also mask finer features of the PMF.\cite{laio2005}
%it also diminishes the eigenvalues
%[by Eq. \eqref{eq:lambda_Gaussian}]
%of more middle-range modes
%(which often represent finer features of the PMF)
%and hence slows down the error reduction there
%[by Eq. \eqref{eq:vt_diffeq_mbin}].
%
Thus, there is an optimal width that balances the two factors.
%
According to the analytical prediction,
the optimal width (assuming the optimal schedule)
was $\sigma = 0.29$, which we used in the simulations,
and it would decrease with the simulation length
and approach the limit of the single-bin updating scheme,
$\sigma \to 0$, asymptotically.

>>>>

\begin{table}[h]
  \caption{\label{tab:lj_error}
    Errors of the bias potential under
    the single-bin,
    Gaussian (with width $\sigma = 0.28$),
    and homogeneous bandpass (with mode cutoff $K = 26$)
    updating schemes
    on the LJ system.
    %
    The numbers have been averaged over multiple runs,
    and the predicted values are shown in parentheses for comparison;
    $\Err_\mathrm{init.}$ and $\Err_\mathrm{final}$
    denote the errors at the beginning and end of the production phase, respectively.
    %
    \note{The reference values can be found
    on the second line of \texttt{verr.log}.
    To check use \texttt{getav.py} under \texttt{prog/lj},
    which will show the simulation and prediction values.}%
  }
  \setlength{\tabcolsep}{2pt}
  \renewcommand\arraystretch{1.4}
  \begin{tabular} { l c c c c }
    \hline
    Scheme & Schedule &
    $\Err_\mathrm{init.}$ &
    $\Err_\mathrm{final}$
    \\
    %\hline
    %\multicolumn{4}{c}{
    %  $\varrho = 0.1$,
    %  $500$ bins,
    %  $\Gamma_1 \approx 111$,
    %  $\Gamma \approx 1.1\times10^3$.
    %} \\
    \hline
    Single-bin & $1/t$
    & $5.7(4.5)\times10^{-2}$
    & $9.1(9.1)\times10^{-6}$
    \\
    %\hline
    Gaussian & $1/t$
    & $6.5(6.2)\times10^{-3}$
    & $1.4(1.4)\times10^{-5}$
    \\
    %\hline
    Gaussian & optimal
    & $6.0(6.2)\times10^{-3}$
    & $2.7(2.7)\times10^{-6}$
    \\
    %\hline
    Bandpass & $1/t$
    & $8.0(7.9)\times10^{-3}$
    & $1.7(1.7)\times10^{-6}$
    \\
    \hline
  \end{tabular}
\end{table}

As shown in Fig. \ref{fig:lj_alpha},
the optimal schedule for the Gaussian updating scheme
with width $\sigma = 0.28$
was more complex than the inverse-time one.
%
We chose the Gaussian updating scheme here because
the optimal and inverse-time schedules were identical
in the single-bin and bandpass updating schemes.
%
The optimal schedule lay above the inverse-time schedule,
and its initial decay was roughly exponential,
in agreement with the discussion in Appendix \ref{sec:schedule_geometry}.
%
Since our equilibration phase was sufficiently long,
the initial updating magnitude, $\alpha(0)$,
was around the ideal value of $a_0/2 = 5\times10^{-5}$
[cf. Eq. \eqref{eq:half_alpha0}].

@@

For the Gaussian updating scheme,
the error depended not only on the schedule
but also on the width of the Gaussian.
%
Although widening the Gaussian
can suppress local noise,
it can also mask finer features of the PMF.\cite{laio2005}
%it also diminishes the eigenvalues
%[by Eq. \eqref{eq:lambda_Gaussian}]
%of more middle-range modes
%(which often represent finer features of the PMF)
%and hence slows down the error reduction there
%[by Eq. \eqref{eq:vt_diffeq_mbin}].
%
Thus, there is an optimal width that balances the two factors.
%
According to the analytical prediction,
the optimal width (assuming the optimal schedule)
was $\sigma = 0.28$, which we used in the simulations,
and it would decrease with the simulation length
and approach the limit of the single-bin updating scheme,
$\sigma \to 0$, asymptotically.
